---
title: "R Notebook"
output: html_notebook
---

# EDA of the Econonomic Policy Uncertainty (EPU) Index

### Goals

* Analyze the properties of the index

### Comments


* EPU Index
  * Source: for Canada: http://www.policyuncertainty.com/canada_monthly.html
  * Includes 5 Canadian Newspapers: The Gazette, The Vancouver Sun, The Toronto Star, The Ottawa Citizen, and The Globe and Mail, Canadian Newswire.
  * Selection of Articles: Utilize the number of news articles containing the terms uncertain or uncertainty, economic or economy, as well as policy relevant terms (scaled by the smoothed number of articles containing 'today')
    * Policy relevant terms include: 'policy', 'tax', 'spending', 'regulation', 'central bank', 'budget', and 'deficit'.
  * Index calculation: 
    * Each paper-specific monthly series is normalized to standard deviation 1 prior to 2011 and then summed across papers, by month. Finally, the series is normalized to mean 100 prior to 2011.
  * data from the preceding two months may be revised slightly: This is driven by the fact that some online newspapers do not immediately update their online archives with all articles, leading to slightly changing totals for the previous 1-2 months.

* Questions
  * Are there any time series breaks? How "reliable" is the index?

```{r}
library("readxl")
library("scales")
library(zoo)
library(ggplot2)
library(plotly)
library(quantmod)
library("xts")
```

## Step 1: Data Retrieval

Downloaded from http://www.policyuncertainty.com/canada_monthly.html

```{r}
epu <- read_excel("/home/jovyan/data/Canada_Policy_Uncertainty_Data.xlsx",
  sheet = 1, col_types = c("text", "text", "numeric")
)
head(epu)
```

Note there is other data in teh same sheet `my_data2 <- read_excel("Canada_Policy_Uncertainty_Data.xlsx",sheet=2)#, `.

## Step 2: Meet and Greet Data


```{r}
epu$date <- paste(epu$Year, epu$Month, sep = "-")
head(epu$date)
```


```{r}
# format date column
epu$date <- as.yearmon(epu$date, format = "%Y-%m")
epu <- epu[-nrow(epu), ]
colnames(epu)[3] <- "epu"
head(epu)
```

```{r}
set.seed(1)
epu %>% sample_n(10)
```

Check for missing values.

Replace character columns with empty string with NA
```{r}
epu <- epu %>% mutate_if(is.character, list(~ na_if(., "")))
```

```{r}
sum(is.na(vectors))
```

Is there missing data in terms of a continous monthly series?

```{r}
print(min(epu$date))
print(max(epu$date))
```

```{r}
complete_sequence <- seq(
  from = as.Date(min(epu$date)),
  to = as.Date(max(epu$date)), by = "month"
)
complete_sequence <- as.yearmon(complete_sequence)
complete_sequence[1:5]
```

There are no missing values which shows the time series does not have missing monthly values.
```{r}
all(epu$date == complete_sequence)
```

No duplicate datapoints found.
```{r}
sum(duplicated(vectors))
```

## Step 3: Data Distributions

```{r}
summary(epu)
```

```{r, fig.width=8, fig.height=4}
ggplot(data = epu, aes(x = epu)) +
  geom_histogram() +
  xlab("epu values")
```


## Step 4: Relationship between variables

Here in particular the time-dependencies of the economic variables are investigated.



```{r}
epu <- xts(df_epu$epu, df_epu$date)
```


label variable allows to show time in tooltip

```{r}
fig <- ggplot(epu, aes(x = Index, y = epu)) +
  geom_line()
ggplotly(fig)
```


```{r}
fig <- ggplot(epu, aes(x = Index, y = epu)) +
  geom_line() +
  xlim(2000, 2018)
fig
```
