---
title: "R Notebook"
output: html_notebook
---

# Investment in Building Construction EDA

### Goals

* Exploration of the Investment in Buildign Construction Table
* Identify 

### Comments

```{r}
suppressMessages(library(cansim))
suppressMessages(library(stringr))
suppressMessages(library(ggplot2))
suppressMessages(library(zoo))
suppressMessages(library(dplyr))
suppressMessages(library(plotly))
```

```{r}
options(cansim.cache_path = "/home/jovyan/data")
```

# Step I: Data Extraction

Investment in building construction table [here](https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=3410017501&pickMembers%5B0%5D=4.1&cubeTimeFrame.startMonth=04&cubeTimeFrame.startYear=2010&cubeTimeFrame.endMonth=08&cubeTimeFrame.endYear=2020&referencePeriods=20100401%2C20200801)

```{r results='hide'}
investment_table <- get_cansim("34-10-0175-01")
```

# Step II: Meet and greet the data

```{r}
head(investment_table)
```

```{r}
nrow(investment_table)
```


rename columns for conformity
```{r}
colnames(investment_table) <- str_replace_all(
    colnames(investment_table), c(" " = "_", "\\(" = "_", "\\)" = "_"))
  colnames(investment_table) <- sapply(colnames(investment_table),
                                            tolower)
```

```{r}
colnames(investment_table)
```


Subselect the relevant columns for further analysis:
```{r}
relevant_columns <- c("type_of_structure", "type_of_work", "investment_value",
                      "vector", "coordinate", "geo", "ref_date",
                      "value",  "uom")
investment <- investment_table[, relevant_columns]
head(investment)
```

### Missing values

```{r}
sapply(investment, function(x) sum(is.na(x)))
```

```{r}
nrow(investment)
```


```{r}
missing_values <- investment[is.na(investment$value), ]
```

```{r}
head(missing_values)
```

```{r}
set.seed(123)
dplyr::sample_n(missing_values, 10)
```


# Step III: Univariate Distributions

### Type of Structure

Important categories: "Total residential and non-residential", 
"Total residential", "Total commercial"
```{r}
unique(investment$type_of_structure)
```


```{r}
table(investment$type_of_structure)
```

All 3 important time series have the same length.
```{r}
investment %>% 
  select(type_of_structure) %>%
  filter(type_of_structure %in% c("Total residential and non-residential",
                        "Total residential", "Total commercial")) %>%
  table()
```

### Type of Work

Important category: "Types of work, total"
```{r}
unique(investment$type_of_work)
```

```{r}
table(investment$type_of_work)
```

```{r}
lapply(table(investment$type_of_work),
       function(x) x / as.numeric(nrow(investment)))
```

### Investment Value

Distinguishes between seasonal adjusted and unajusted, constant and current. 
On the StatCan website defines the difference:
"Deflation: The methodology of transforming current dollar values to constant dollar values (deflation process) has been updated from 2007=100 to 2012=100." 
Steve's comment is to use current dollars for now as we can transform them
into constant dollars later.


```{r}
unique(investment$investment_value)
```

```{r}
lapply(table(investment$investment_value),
       function(x) x / as.numeric(nrow(investment)))
```


### Time Span

Time span available is from 1/2010 to 8/2020. 8/2020 is expected due to the publication lag of two months. Same number of data points per month for the whole time span.

```{r}
table(investment$ref_date)
```

```{r}
unique(table(investment$ref_date))
```

### Geographic Information

Geographic information is avialable on the country level ("Canada"), on the province level and on some cases, the city areas.
```{r}
unique(investment$geo)
```

Available provinces extracted from the above list:
```{r}
province_names <- c("British Columbia", "Nova Scotia", "Nunavut", "Quebec",
                    "Saskatchewan",
                    "Alberta", "Manitoba", "New Brunswick",
                    "Newfoundland and Labrador", "Northwest Territories",
                    "Ontario", "Prince Edward Island", "Yukon")
length(province_names)
```

> All 13 provinces and territories available.

### Units

All data is in dollars.
```{r}
unique(investment$uom)
```

### Vector and Coordinates

Both are used to subset the data to create series of data. Vectors allow to choose the right time series for our purposes

```{r}
length(unique(investment$vector))
```

## Step IV: Multivariate Analysis

Subselect relevant data and compare time series by categories.


### Total 

```{r}
total <- investment %>%
  filter(type_of_structure == "Total residential and non-residential") %>%
    filter(type_of_work == "Types of work, total") %>%
      filter(investment_value == "Unadjusted - current") %>%
        filter(geo == "Canada")
```

```{r}
nrow(total)
```

time series `total` corresponds to one datapoint per month.
```{r}
lapply(total, unique)
```


No missing values.
```{r}
sum(is.na(total))
```


Findings of total time series
* Time series reveals clear trend and seasonal effects. 
* Extreme drop due to Covid-19.
* Peaks pronounced in July/August, with a second weak peak around November
* The august value is clearly higher than what one can infer from the trend 
of previous years


```{r}
total$type_of_structure <- "Total"
```




```{r}
p <- ggplot(data=total, aes(x=as.yearmon(ref_date), y=value)) + geom_line() +
              labs(title = "Total Investment", 
                   x = "reference month", y = "value (dollars)")
ggplotly(p)
```


### Comparing Residential and Non-Residential

```{r}
total_res_com <- investment %>%
  filter(type_of_structure %in% c("Total residential", "Total commercial")) %>%
    filter(type_of_work == "Types of work, total") %>%
      filter(investment_value == "Unadjusted - current") %>%
        filter(geo == "Canada")
```


```{r}
lapply(total_res_com, unique)[c(1,2,3,4,5,6)]
```

```{r}
p <- ggplot(data=total_res_com) + geom_line(aes(x=as.yearmon(ref_date), y=value, 
                      group=type_of_structure, linetype=type_of_structure)) +
              geom_line(data=total, aes(x=as.yearmon(ref_date), y=value, linetype=type_of_structure, alpha=0.1)) +
              labs(title = "Comparison Residential - Commerical", x = "reference month", y = "value (dollars)",
                   alpha= "", linetype= "Type of Structure")
ggplotly(p, width=800)
```



### Comparison Total by Geography

```{r}
total_by_geo <- investment %>%
  filter(type_of_structure == "Total residential and non-residential") %>%
    filter(type_of_work == "Types of work, total") %>%
      filter(investment_value == "Unadjusted - current") %>%
        filter(geo %in% province_names)
```

```{r}
lapply(total_by_geo, unique)[c(1,2,3,4,5,6)]
```

* Ontario has by far the most investment, compared to other provinces
* 

```{r, fig.height=7}
p <- ggplot(data=total_by_geo) + geom_line(aes(x=as.yearmon(ref_date),
                        y=value, group=geo, color=geo)) +
            geom_line(data=total, aes(x=as.yearmon(ref_date), y=value, 
                                      color=geo, alpha=0.1)) +
              labs(title = "Comparison Total by Geography",
                   x = "reference month", y = "value (dollars)", 
                            alpha= "", color= "Province/Country") +
            scale_y_continuous(trans='log10')
ggplotly(p, width=800, height=490)
```




