---
title: "ValidationSchemesTest"
output: html_notebook
---

### Goals

* Develop validation methods for training, validation and test phases.
* Use building permits for testing of setup to ensure as close as possible to
real ase.

### Comments

* Get clarification on type of work and other variables to choose correct
series.

```{r, warning=FALSE, message=FALSE, include=FALSE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
suppressMessages(library(reticulate))
suppressMessages(library(rjson))
library(fpp3)
library(tsibble)
import("datetime")
import("elasticsearch")
import("certifi")

library(tidyverse)
```

# I. Retrieve data

Load the data.
```{r}
path <- "/home/jovyan/forecasting_data_ingestion/notebooks/"
source_python(paste(path, "load_es_data.py", sep=""))
creds = fromJSON(file=paste(path, "cred.json", sep=""))

indexname = "buildingpermits"
permits <- load_from_es(creds, indexname)
```

```{r}
dim(permits)
```

```{r}
head(permits)
```


```{r}
permits$ref_date <- as.Date(permits$ref_date)
```


# II. Data Exploration 


* There are different categories of series, 

Building permits start at 2011!
```{r}
summary(permits)
```

```{r}
min(as.Date(permits$ref_date))
```


## Different Categories of series

Pick value of permits, choose `Value of permits `. Which one to forecast?

### variable

```{r}
table(permits$variables)
```

### type_of_structure

Choose `Total residential and non-residential` for now.

What represents the non-residential sector: `Total non-residential` 
(or `Total industrial` or ` Total commercial`)

```{r}
table(permits$type_of_structure)
```

### Type of work

Choose `Types of work, total ` for now. Are we only interested
in the total `Types of work, total` or in `New construction`.
```{r}
table(permits$type_of_work)
```

### Geo Level

Choose Canada
```{r}
table(permits$geo)
```

### Adjustment

```{r}
table(permits$seasonal_adjustment)
```

## Series of interest overview


```{r}
permits_series_set <- permits %>%
  filter(variables=="Value of permits", 
         (type_of_structure == "Total residential and non-residential" |
          type_of_structure == "Total non-residential" | 
          type_of_structure == "Total residential"),
         type_of_work=="Types of work, total",
         #geo=="Canada", all
         seasonal_adjustment=="Seasonally adjusted")
permits_series_set <- permits_series_set[, c("ref_date", "type_of_structure", "geo", "value")]
head(permits_series_set)
```

```{r}
permits_series_cols <- pivot_wider(permits_series_set, names_from = c("type_of_structure", "geo"), values_from=value)
colnames(permits_series_cols) <- str_replace_all( 
  colnames(permits_series_cols), c(" " = "_", "\\(" = "_", "\\)" = "_"))
head(permits_series_cols)
```




```{r, fig.width=12, fig.height=12}
ggplot(permits_series_set) + geom_line(aes(x=ref_date, y=value, 
                                           color = type_of_structure)) +
  facet_wrap( ~ geo, nrow = 12, scales = "free_y") +
  labs(title = "Permit Value")
```

### Missing Values

Above series indicates ther is missing data for Nunavut and Northwestern 
Territories. Below I show which exact dates and which values are missing.

```{r}
permits_series_ts <-permits_series_cols %>% 
  mutate(date = yearmonth(ref_date)) %>%
    as_tsibble(index = date) %>%
      subset(select=-c(ref_date))
head(permits_series_ts)
```


```{r}
# we explicitly fill month gaps with nan, if a month data is missing
explicit_missingness <- fill_gaps(permits_series_ts) # %>% filter(is.na(value))

explicit_missingness <- explicit_missingness[rowSums(is.na(explicit_missingness)) > 0,]
cbind(explicit_missingness["date"], explicit_missingness[sapply(explicit_missingness, function (x) sum(is.na(x)) >0)])

```




## Subselect on series for analysis

```{r}
permits_series <- permits %>%
  filter(variables=="Value of permits", 
         type_of_structure"Total residential and non-residential",
         type_of_work=="Types of work, total",
         geo=="Canada",
         seasonal_adjustment=="Seasonally adjusted")
head(permits_series)
```

Ensure unique series after filtering.
```{r}
sapply(permits_series, function (x) length(unique(x)))
dim(permits_series)
```

## Visualization of test series

```{r}
ggplot(permits_series) + geom_line(aes(x=as.Date(ref_date), y=value))
```

