---
title: "TestSetup"
output: html_notebook
# output: rmarkdown::github_document
# output: 
#  html_document:
#    keep_md: true
---

# Exploratory analysis of Retail Trade Sales

### Goals

* Reproduce the Index of the Canadian Economic Dashboard and COVID-19
* Identify data tables and vectors for the dashboard

### Comments

* Time series from Canadian Economic Dashboard and COVID-19 at https://www150.statcan.gc.ca/n1/pub/71-607-x/71-607-x2020009-eng.htm
* There are 16 indicators


```{r}
suppressMessages(library(cansim))
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(plotly))
source("../../src/utils.R")
```

## Extract the Data
```{r}
vector_names <- c(
  "v65201210", "v41690973", "v2062809", "v1001827265",
  "v1001826653", "v52367097", "v4391505", "v800450",
  "v32858858", "v32858865", "v32858872", "v74869", "v129449",
  "v129460", "v129472", "v129482"
)
length(vector_names)
```

```{r}
vector_descr <- c(
  "RealGDP", "CPI", "Employment [pers]", "Int. merchendise trade Exp. [$]",
  "Int. merchendise trade Imp. [$]", "Retail Sales [$]", "hours worked",
  "Manufact. sales [$]", "Aircraft domestic [#]", "Aircraft transborder [#]",
  "Aircraft int other [#]", "Railway carloads [tons]", "Travelers US [pers]",
  "Travelers other country [pers]", "CA resident US [pers]",
  "CAresident other country [pers]"
)
names(vector_descr) <- vector_names
```

I expect this date to be before the start of all time series/vectors.
```{r}
start_date <- "1900-01-01"
```

Retrieve all vectors. They are concatunated along axis 0.
```{r}
vectors <- get_cansim_vector(vector_names, start_date)
dim(vectors)
```


## Preprocessing

Normalization all indicators so that December 2019 = 100

```{r}
colnames(vectors) <- sapply(colnames(vectors), tolower)
colnames(vectors)
```

```{r}
vectors$ref_date <- as.Date(vectors$ref_date)
```


```{r}
head(vectors)
```

```{r}
vector_names[1]
```

```{r}
length(vector_names)
```


```{r}
indicators1 <- vector(length = 16)
for (i in 1:seq_len(vector_names)) {
  indicator <- filter(vectors, vector == vector_names[i])
  ref_value <- indicator$value[indicator$ref_date == as.Date("2019-12-01")]
  indicator$value_index <- 100. * indicator$value / ref_value
  if (i == 1) {
    indicators <- indicator
    }
  else {
    indicators <- rbind(indicators, indicator)
    }
}
```

```{r}
head(indicators)
```

```{r}
indicators$descr <- sapply(indicators$vector, function(x) {
  vector_descr[x][[1]]
})
```




## Canadian Economic Dashboard Visualization

```{r}
p <- ggplot(indicators, aes(x = ref_date, y = value_index)) +
  geom_line((aes(group = descr, color = descr))) +
  scale_x_date(date_breaks = "2 month", date_labels = "%b %Y",
               limits = as.Date(c("2019-01-01", "2020-06-01"))) +
  scale_y_continuous(breaks = seq(0, 180, 20), limits = c(0, 180)) +
  labs(title = "Index, December 2019 = 100", x = "", y = "")
ggplotly(p)
```

