---
title: "I. Extraction of Dashboard Vectors"
output: html_notebook
---

### Goals

* Identification of variables/vectors from Covid-19 dashboard in the STC tables.

### Comments


```{r}
suppressMessages(library(cansim))
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(plotly))
suppressMessages(library(stringr))
```



cache tables in data directory
```{r}
options(cansim.cache_path = "/home/jovyan/data")
```


## A. Data Extraction from API

As defined under Definitions section at 
https://www150.statcan.gc.ca/n1/pub/71-607-x/71-607-x2020009-eng.htm. 
Note that a snapshot of the website state at time of development of 
this notebook is in ../references/Canadian Economic Dashboard and COVID-19.pdf
```{r}
dashboard_table_names <- c(
  "36-10-0434-01", "18-10-0004-01", "14-10-0287-01", "12-10-0121-01",
  "12-10-0121-01",
  "20-10-0008-01", "14-10-0289-01", "16-10-0047-01", "23-10-0008-01",
  "23-10-0008-01",
  "23-10-0008-01", "23-10-0216-01", "24-10-0005-01", "24-10-0005-01",
  "24-10-0005-01",
  "24-10-0005-01"
)
length(dashboard_table_names)
```

Download all NDM tables (uncomment when need to download again)
```{r}
dashboard_tables <- lapply(dashboard_table_names, FUN = get_cansim)
```

## B. Preprocessing and Vector Identification

Replace spaces in variable names with underscore for better processing. Search through tables and identify correct vectors for dashboard. List of vectors in D.
```{r}
for (i in seq_along(dashboard_tables)) {
  colnames(dashboard_tables[[i]]) <- str_replace_all(
    colnames(dashboard_tables[[i]]), c(" " = "_", "\\(" = "_", "\\)" = "_")
  )
  colnames(dashboard_tables[[i]]) <- sapply(
    colnames(dashboard_tables[[i]]),
    tolower
  )
}
```



```{r}
head(filter(
  dashboard_tables[[4]], trade == "Export",
  seasonal_adjustment == "Seasonally adjusted",
  basis == "Balance of payments"
))
```

Same table for exports and the imports here
```{r}
head(filter(
  dashboard_tables[[4]], trade == "Import",
  seasonal_adjustment == "Seasonally adjusted",
  basis == "Balance of payments"
))
```

```{r}
head(filter(dashboard_tables[[6]], adjustments == "Seasonally adjusted"))
```

```{r}
dashboard_table_names[[8]]
```


```{r}
head(filter(
  dashboard_tables[[8]], seasonal_adjustment == "Seasonally adjusted",
  principal_statistics == "Sales of goods manufactured (shipments)",
  north_american_industry_classification_system__naics_ ==
    "Manufacturing [31-33]"
))
```

```{r}
head(filter(
  dashboard_tables[[9]], type_of_operation ==
    "Total itinerant movements",
  domestic_and_international_itinerant_movements ==
    "Domestic movements"
))
```

same table as 9
```{r}
head(filter(
  dashboard_tables[[10]], type_of_operation ==
    "Total itinerant movements",
  domestic_and_international_itinerant_movements ==
    "Transborder movements"
))
```

same table as 9
```{r}
head(filter(
  dashboard_tables[[11]],
  type_of_operation == "Total itinerant movements",
  domestic_and_international_itinerant_movements ==
    "Other international movements"
))
```

```{r}
head(filter(dashboard_tables[[14]]))
```

```{r}
head(filter(dashboard_tables[[14]], traveller_category ==
  "Travellers from countries other than United States"))
```

```{r}
head(filter(dashboard_tables[[15]], traveller_category ==
  "Canadian travellers returning from United States"))
```

```{r}
head(filter(dashboard_tables[[16]], traveller_category ==
  "Canadian travellers returning from countries other than United States"))
```



## C) Dashboard vectors Identified

Vectors identified through the previous analysis. Order 
```{r}
vectors <- c(
  "v65201210", "v41690973", "v2062809", "v1001827265",
  "v1001826653", "v52367097", "v4391505", "v800450",
  "v32858858", "v32858865", "v32858872", "v74869",
  "v129449", "v129460", "v129472", "v129482"
)
length(vectors)
```

Notes to the selection of vectors:
  
* GDP: all industries
* CPI: All-items consumer price index (2002=100), not adjusted for seasonality. Note that UOM is set as 2002=100.
* exports (4): assume `Total of all merchandise` is equal to `total of all products` in Definitions on website, filtered by `Export` and `Seasonal adjustment`, 'Balance of Payments`. Assume uOM dollars is current dollars.
* Manufacturers' sales: assume `current dollars` is `dollars`, assume choice is `Manufacturing [31-33]` as NAICS, as other NAICS are under this category.
* Travellers from the United States: assume `Traveller_category`=`United States travellers` is correct.
