---
title: "R Notebook"
output: html_notebook
---

# List of TSX Issuers

### Goals

* Analyse composition of stocks from TSX Issuers related to Construction and Real Estate

### Notes

* Data directly from TSX at https://www.tsx.com/listings/current-market-statistics
and specifically. This data we would want to ingest at some point.

```{r}
rm(list=ls())
```


```{r}
library("readxl")
library(quantmod)
library(dplyr)
library(ggplot2)
```

## I. Loading and Preprocessing of Excel file


The file below was downloaded 26/10/2020 from the TSX website above.
File contains first some rows with non-relevant information to skip
```{r}
issuers <- read_excel(paste("../../../data_tmp/tsx-andamp-tsxv-listed", 
                            "-companies-2020-10-16-en-1.xlsx", sep=""), 
                      sheet=2, skip=6)
```

```{r}
head(issuers, 2)
```

```{r}
colnames(issuers) <- lapply(str_replace_all(colnames(issuers), "[\r\n -$\\(\\)\\-]", "_"), FUN=tolower)
```


Discard companies from the sector `Real Estate Operating Companies` as not
relevant based on Elton's comments.
```{r}
issuers <- issuers %>% filter(real_estate_sub_sector != 
                      "Real Estate Operating Companies")
```

```{r}
real_estate <- filter(issuers, !is.na(issuers[, 'real_estate_sub_sector']))[,  c("real_estate_sub_sector", "name", "root__ticker", "market_cap__c____30_september_2020")]
head(real_estate)
```


### Download of Stocks based on Tickers

Root ticker are not representative of the TSX

```{r}
head(real_estate)  
```


Use `ticker` column to download data. Try statement to catch errors.

```{r}
#stock_name <- "BPY-UN.TO"

get_stockdata <- function(stock_name_root) {

  out <- try ({
    stock_name <- paste(stock_name_root, "-UN.TO", sep="")
    stock <- getSymbols(stock_name, src = "yahoo", auto.assign = FALSE) 

    # calculate monthly returns instead of using all stock data
    stock = monthlyReturn(stock)
    colnames(stock) <- sapply(colnames(stock), function (colstock) str_split(colstock, "\\.")[[1]][length(str_split(colstock, "\\.")[[1]])])
    stock <- data.frame(date=index(stock), coredata(stock))
    stock$stock_name <- rep(stock_name, nrow(stock))
    stock$stock_name_root <- rep(stock_name_root, nrow(stock))
    stock
  })
  
  # if stock is not available as "[stockname]-UN.TO", try "[stockname"].TO"
  if (class(out) == "try-error") {
  
    out <- try ({
      stock_name <- paste(stock_name_root, ".TO", sep="")
      stock <- getSymbols(stock_name, src = "yahoo", auto.assign = FALSE) 
    # calculate monthly returns instead of using all stock data
      stock = monthlyReturn(stock)
      colnames(stock) <- sapply(colnames(stock), function (colstock) str_split(colstock, "\\.")[[1]][length(str_split(colstock, "\\.")[[1]])])
      stock <- data.frame(date=index(stock), coredata(stock))
      stock$stock_name <- rep(stock_name, nrow(stock))
      stock$stock_name_root <- rep(stock_name_root, nrow(stock))
      stock
    })
  }
  
  out
  
  }
```


```{r}
start.time <- Sys.time()
get_stock_series <- lapply(real_estate$root__ticker, get_stockdata)
print(paste("total download time :", Sys.time() - start.time, "s"))
```


Identify which time series was not downloaded successfully and print their 
error messages. Successfull stocks are saved in stock_series.
```{r}
downloaded <- list()
stock_series <- list()
j = 1
k = 1
for (i in 1:length(get_stock_series)) {
  if (class(get_stock_series[[i]]) == "try-error") {
    #missing_data_log[[k]] <- get_stock_series[[i]]
    print(get_stock_series[[i]])
    downloaded[[i]] <- FALSE
  } else {
    downloaded[[i]] <- TRUE
    stock_series[[j]] <- get_stock_series[[i]]
    j <- j + 1
  }
}
real_estate['downloaded'] <- unlist(list(downloaded))
```

### Analysis of missing data

A small amount of stocks were not downloaded/are not available
```{r}
table(real_estate$downloaded)
```

The following 3 stocks could not be downloaded. In fact 
* BPO.TO: shows NA values in Yahoo: https://ca.finance.yahoo.com/quote/bpo.to/
* ERE: not available on Yahoo, should be available on TSX, according 
to https://www.eresreit.com. As its European I discard the ticker for now.
* SGR: should be availalbe at https://ca.finance.yahoo.com/quote/SGR-UN.TO/
Note that all 3 missing stocks are part of the 
`Industrial/Office/Retail/Residential` subsector for which plenty of companies
are available already. Discard the companies below in further work.
```{r}
real_estate[ !real_estate$downloaded, ]
```



## II. Analysis of stocks Properties

```{r}
stock_series <- bind_rows(stock_series)
```



```{r}
head(stock_series)
```

add sub-sector information to stocks
```{r}
stock_series <- merge(stock_series, real_estate, by.x="stock_name_root", by.y="root__ticker", all=FALSE)
head(stock_series, 2)
```


```{r}
table(real_estate[real_estate$downloaded, ]$real_estate_sub_sector)
```

### 1. Category: Diversified

Most companies go back to 2007 but those with the largest market cap
go back only to 2012. Fair amount of distribution.
```{r}
stock_series_stats_div <- stock_series %>% 
  filter(real_estate_sub_sector=="Diversified") %>%
  group_by(stock_name_root) %>%
      summarize(
        min = min(date, na.rm=TRUE),
        max = max(date, na.rm=TRUE),
        real_estate_sub_sector = unique(real_estate_sub_sector),
        market_cap__c____30_september_2020 = 
          max(market_cap__c____30_september_2020),
        name = unique(name),
        .groups = "keep"
      )
stock_series_stats_div$market_cap_fraction <- 
  stock_series_stats_div$market_cap__c____30_september_2020/sum(stock_series_stats_div$market_cap__c____30_september_2020)
stock_series_stats_div$real_estate_sub_sector <- NULL
stock_series_stats_div
```

```{r, fig.width=24, fig.height=12}
p <- ggplot(data = filter(stock_series, real_estate_sub_sector == "Diversified")) + 
              geom_line(aes(x = as.Date(date), y = returns, group=stock_name_root, color=name, linetype=name)) +
                labs(title = "Category - Diversified",
                   x = "reference month", y = "monthly stock returns", 
                            group= "Ticker", color= "Ticker")
ggplotly(p, width=880, height=450)
```


### 2. Category: Specialized

The companies with a largest market capitalization exist also the longest.
If we would weight according to capitalization, then only two companies would
dominate an categorical indicator.
```{r}
stock_series_stats_spec <- stock_series %>% 
  filter(real_estate_sub_sector=="Specialized") %>%
  group_by(stock_name_root) %>%
      summarize(
        min = min(date, na.rm=TRUE),
        max = max(date, na.rm=TRUE),
        real_estate_sub_sector = unique(real_estate_sub_sector),
        market_cap__c____30_september_2020 = 
          max(market_cap__c____30_september_2020),
        name = unique(name),
        .groups = "keep"
      )
stock_series_stats_spec$market_cap_fraction <- 
  stock_series_stats_spec$market_cap__c____30_september_2020/sum(stock_series_stats_spec$market_cap__c____30_september_2020)
stock_series_stats_spec$real_estate_sub_sector <- NULL
stock_series_stats_spec
```

```{r, fig.width=24, fig.height=16}
p <- ggplot(data = filter(stock_series, real_estate_sub_sector == "Specialized")) + 
              geom_line(aes(x = as.Date(date), y = returns, group=stock_name_root, color=name, linetype=name)) +
                labs(title = "Category - Specialized",
                   x = "reference month", y = "close stock value", 
                            group= "Ticker", color= "Ticker")
ggplotly(p, width=80, height=450)
```


### 3. Category: Real Estate Development and Services

All 3 companies exist for a longer time at the stock market and have a similar
market cap. One company would dominate a categorical indicator with 50%.

```{r}
stock_series_stats_real <- stock_series %>% 
  filter(real_estate_sub_sector=="Real Estate Development and Services") %>%
  group_by(stock_name_root) %>%
      summarize(
        min = min(date, na.rm=TRUE),
        max = max(date, na.rm=TRUE),
        real_estate_sub_sector = unique(real_estate_sub_sector),
        market_cap__c____30_september_2020 = 
          max(market_cap__c____30_september_2020),
        name = unique(name),
        .groups = "keep"
      )
stock_series_stats_real$market_cap_fraction <- 
  stock_series_stats_real$market_cap__c____30_september_2020/sum(stock_series_stats_real$market_cap__c____30_september_2020)
stock_series_stats_real$real_estate_sub_sector <- NULL
stock_series_stats_real
```

```{r, fig.width=24, fig.height=14}
p <- ggplot(data = filter(stock_series, real_estate_sub_sector == "Real Estate Development and Services")) + 
              geom_line(aes(x = as.Date(date), y = returns, group=stock_name_root, color=name, linetype=name)) +
                labs(title = "Category - Real Estate Development and Services",
                   x = "reference month", y = "close stock value", 
                            group= "Ticker", color= "Ticker")
ggplotly(p, width=780, height=450)
```


### 4. Category: Industrial/Office/Retail/Residential

```{r}
stock_series_stats_ind <- stock_series %>% 
  filter(real_estate_sub_sector=="Industrial/Office/Retail/Residential") %>%
  group_by(stock_name_root) %>%
      summarize(
        min = min(date, na.rm=TRUE),
        max = max(date, na.rm=TRUE),
        real_estate_sub_sector = unique(real_estate_sub_sector),
        market_cap__c____30_september_2020 = 
          max(market_cap__c____30_september_2020),
        .groups = "keep"
      )
stock_series_stats_ind$market_cap_fraction <- 
  stock_series_stats_ind$market_cap__c____30_september_2020/sum(stock_series_stats_ind$market_cap__c____30_september_2020)
stock_series_stats_ind$real_estate_sub_sector <- NULL
stock_series_stats_ind
```

Top companies in company for market cap indicates that the market cap is fairly
distributed and there may be a less dependency of a categorical indicator on
individual companies.
```{r}

head(stock_series_stats_ind %>% arrange(across(market_cap_fraction, desc)), 10)
```

```{r, fig.width=24, fig.height=12}
p <- ggplot(data = filter(stock_series, real_estate_sub_sector == "Industrial/Office/Retail/Residential")) + 
              geom_line(aes(x = as.Date(date), y = returns, group=stock_name_root, color=name, linetype=name)) +
                labs(title = "Category - Industrial/Office/Retail/Residential",
                   x = "reference month", y = "close stock value", 
                            group= "Ticker", color= "Ticker")
ggplotly(p, width=880, height=450)
```


## 5. All Categories

```{r}
stock_series_stats_all <- stock_series %>% 
  group_by(stock_name_root) %>%
      summarize(
        min = min(date, na.rm=TRUE),
        max = max(date, na.rm=TRUE),
        real_estate_sub_sector = unique(real_estate_sub_sector),
        market_cap__c____30_september_2020 = 
          max(market_cap__c____30_september_2020),
        .groups = "keep"
      )
stock_series_stats_all$market_cap_fraction <- 
  stock_series_stats_all$market_cap__c____30_september_2020/sum(stock_series_stats_all$market_cap__c____30_september_2020)
stock_series_stats_all$real_estate_sub_sector <- NULL
stock_series_stats_all
```


Fairly well distributed.
```{r}
head(stock_series_stats_all %>% arrange(across(market_cap_fraction, desc)), 10)
```

```{r, fig.width=24, fig.height=12}
p <- ggplot(data = stock_series) + 
              geom_line(aes(x = as.Date(date), y = returns, group=stock_name_root, alpha=0.2)) +
                labs(title = "All tickers",
                   x = "reference month", y = "close stock value", 
                            group= "Ticker", color= "Ticker")
ggplotly(p, width=880, height=450)
```







