---
title: "R Notebook"
output: html_notebook
---

# List of TSX Issuers

### Goals

* Analyse composition of TSX Issuers related to Construction and Real Estate

### Notes

* Data directly from TSX at https://www.tsx.com/listings/current-market-statistics
and specifically. This data we would want to ingest at some point.


```{r}
rm(list=ls())
library("readxl")
```

## I. Loading and Preprocessing of Excel file



The file below was downloaded 26/10/2020 from the TSX website above.
File contains first some rows with non-relevant information to skipo
```{r}
issuers_tsx <- read_excel(paste("../../../data_tmp/tsx-andamp-tsxv-listed", 
                            "-companies-2020-10-16-en-1.xlsx", sep=""), 
                      sheet=2, skip=6)

issuers_tsxv <- read_excel(paste("../../../data_tmp/tsx-andamp-tsxv-listed", 
                            "-companies-2020-10-16-en-1.xlsx", sep=""), 
                      sheet=3, skip=6)

```

```{r}
head(issuers_tsx)
```

```{r}
head(issuers_tsxv)
```

```{r}
colnames(issuers_tsx) <- lapply(str_replace_all(colnames(issuers_tsx), "[\r\n -$\\(\\)\\-]", "_"), FUN=tolower)
colnames(issuers_tsxv) <- lapply(str_replace_all(colnames(issuers_tsxv), "[\r\n -$\\(\\)]", "_"), FUN=tolower)
colnames(issuers_tsx)
```

```{r}
cols_of_interest <- c("name" , "real_estate_sub_sector", 
      "market_cap__c____30_september_2020", "root__ticker")
```

Based on David Bertrand's comments I am not considinger TSVX stocks here.
```{r}
# issuers <- rbind(issuers_tsx[, cols_of_interest], issuers_tsxv[, cols_of_interest])
issuers <- issuers_tsx[, cols_of_interest]
```

## II. Analysis of Real-Estate Companies



```{r}
real_estate <- filter(issuers, !is.na(issuers[, 'real_estate_sub_sector']))
head(real_estate)
```

All issuers are from the TSX.

```{r}
real_estate_companies <- real_estate[, c("real_estate_sub_sector", "name", "root__ticker", "market_cap__c____30_september_2020")]
```

```{r}
nrow(real_estate_companies)
```

```{r, fig.height=7}
ggplot(real_estate_companies) + geom_bar(aes(x=reorder(name, market_cap__c____30_september_2020), y=market_cap__c____30_september_2020, fill=real_estate_sub_sector), 
                                         stat="identity") + coord_flip() + labs(x="Market Cap", y="CAD", title = "Market capitalization")
```


### Analysis of the 5 different real-estate subsectors

```{r}
table(real_estate_companies[,'real_estate_sub_sector'])
```

Industrial/Office/Retail/Residential
```{r}
filter(real_estate_companies, real_estate_sub_sector=="Industrial/Office/Retail/Residential")[, c("name", "root__ticker", "market_cap__c____30_september_2020")]
```

Real Estate Operating Companies: 
Elton mentioned these companies are not of interest.
```{r}
filter(real_estate_companies, real_estate_sub_sector=="Real Estate Operating Companies")[, c("name", "root__ticker", "market_cap__c____30_september_2020")]
```

Real Estate Development and Services: 
```{r}
filter(real_estate_companies, real_estate_sub_sector=="Real Estate Development and Services")[, c("name", "root__ticker", "market_cap__c____30_september_2020")]
```


```{r}
filter(real_estate_companies, real_estate_sub_sector=="Specialized")[, c("name", "root__ticker", "market_cap__c____30_september_2020")]
```

```{r}
filter(real_estate_companies, real_estate_sub_sector=="Specialized")[, c("name", "root__ticker", "market_cap__c____30_september_2020")]
```



