---
title: "R Notebook"
output: html_notebook
---

# List of TSX Issuers

### Goals

* Analyse composition of TSX Issuers related to Construction and Real Estate

### Notes

* File provided by David Bertrand from 2019


```{r}
rm(list=ls())
library("readxl")
```

## I. Loading and Preprocessing of Excel file

```{r}
issuers <- read_excel("../../../data_tmp/tsx_listed_issuers_2019.xlsx")
```

```{r}
head(issuers)
```


```{r}
colnames(issuers) <- lapply(str_replace_all(colnames(issuers), "[\r\n -]", "_"), FUN=tolower)
colnames(issuers)
```

## II. Univariate Analysis

### Sectors

```{r}
table(issuers$sector)
```

### Subsector

```{r}
table(issuers$sub__sector)
```


### Real Estate Subsector

AVailable in real_estate_sub-sector

```{r}
table(issuers[, 'real_estate_sub_sector'])
```


## III. Analysis of Real-Estate Companies



```{r}
real_estate <- filter(issuers, !is.na(issuers[, 'real_estate_sub_sector']))
head(real_estate)
```

All issuers are from the TSX:
```{r}
table(real_estate$exchange)
```

```{r}
real_estate_companies <- real_estate[, c("real_estate_sub_sector", "name", "root__ticker")]
```


### Analysis of the 5 different real-estate subsectors

```{r}
table(real_estate_companies[,'real_estate_sub_sector'])
```

Industrial/Office/Retail/Residential
```{r}
filter(real_estate_companies, real_estate_sub_sector=="Industrial/Office/Retail/Residential")[, c("name", "root__ticker")]
```

Real Estate Operating Companies: 
```{r}
filter(real_estate_companies, real_estate_sub_sector=="Real Estate Operating Companies")[, c("name", "root__ticker")]
```

Real Estate Development and Services: 
```{r}
filter(real_estate_companies, real_estate_sub_sector=="Real Estate Development and Services")[, c("name", "root__ticker")]
```


