---
title: "II. Weather and Climate Data"
output: html_notebook
---

### Goals

* Authomatically downloading weather and climate data

### Comments

* For downloading data from Environment and Climate Change Canada (ECCC) website,
we use package `weathercan` which can be installed directly from CRAN. 

* To downlaod data using this package, we need to know the `station_id`
associated with the station we are interested in.


```{r}
suppressMessages(library(tidyverse))
suppressMessages(library(weathercan))
suppressMessages(library(ggplot2))
```

```{r}
head(stations)
```

```{r}
glimpse(stations)
```

```{r}
str(stations)
```

```{r}
summary(stations)
```

```{r}
stations_search("Toronto City", interval = "day")
```

```{r}
toronto_city <- weather_dl(station_ids = 31688, start = "2020-02-01", end = "2020-04-15", interval = "day")
```

```{r}
dim(toronto_city)
```

```{r}
colnames(toronto_city)
```

```{r}
glimpse(toronto_city)
```

```{r}
summary(toronto_city)
```


```{r}
toronto_city %>%
  with(hist(mean_temp))
```

```{r}
toronto_city %>%
  with(hist(total_precip))
```

```{r}
toronto_city %>%
  with(hist(snow_grnd))
```

```{r}
ggplot(data = toronto, aes(x = date, y = mean_temp, group = station_name, colour = station_name)) +
  theme_minimal() + 
  geom_line()
```


```{r}
ggplot(data = toronto, aes(x = date, y = snow_grnd, group = station_name, colour = station_name)) +
  theme_minimal() + 
  geom_line()
```

```{r}
ggplot(data = toronto, aes(x = date, y = total_precip, group = station_name, colour = station_name)) +
  theme_minimal() + 
  geom_line()
```

### Climate Normals
To access climate normals, we first need to know the `climate_id` associated with the station we are interested in.

```{r}
stations_search("Toronto", normals_only = TRUE)
```

```{r}
toronto_city_normal <- normals_dl("6158733")
```

There are two parts to the normals data, average weather measurements and average frost dates.

```{r}
unnest(toronto_city_normal, normals)
```

```{r}
unnest(toronto_city_normal, frost)
```