---
title: "R Notebook"
output: html_notebook
---

# Extraction of Employment Variables

### Goals

* Identify necessary employment-related variables
* Identify vectors related to these variables

### Comments

* Unemployment - Employment relation
  * The unemployment rate is the number of unemployed persons expressed as a 
  percentage of the labour force.
  * The labour force are estimates are provided every month
* LFS data are used to produce the well-known unemployment rate
* Hence forecasting the unemployment rate will not be sufficient to derive the
employment in person as in the economic indicator. One would also approximate
the labour force?
* employment table

```{r}
library(cansim)
library(stringr)
```

## Identification of employment-related data vectors

```{r}
dashboard_table <- get_cansim('14-10-0287-01')
```

Replace spaces in variable names with underscore for better processing. Search through tables and identify correct vectors for dashboard. List of vectors in D.
```{r}

colnames(dashboard_table) <- str_replace_all(colnames(dashboard_table), 
                                             c(" " = "_", 
                                               "\\(" = "_", "\\)" = "_"))
```


```{r}
head(dashboard_table)
```

```{r}
table(dashboard_table$Labour_force_characteristics)
```

```{r}
colnames(dashboard_table)
```

```{r}
class(dashboard_table)
```

### Retrieving the labour force vector.
```{r}
head(dashboard_table[
                  (dashboard_table$Labour_force_characteristics=='Labour force') 
                & (dashboard_table$Data_type=='Seasonally adjusted')
                & (dashboard_table$Age_group=='15 years and over'),])
                #& (dashboard_table$Statistics=='Estimate'),])
```

Estimate of the labour force vector is `v2062810`. There is also a 
standard error of this estimate which is the vector `v101884808`.

### Retrieving the number of unemployed persons
```{r}
head(dashboard_table[
                  (dashboard_table$Labour_force_characteristics=='Unemployment') 
                & (dashboard_table$Data_type=='Seasonally adjusted')
                & (dashboard_table$Age_group=='15 years and over'),])
```

The estimate of the unemployed persons is `v2062814`. The standard error is
`v101885096`.

### Retrieving the unemployment rate
```{r}
head(dashboard_table[
                  (dashboard_table$Labour_force_characteristics==
                     'Unemployment rate') 
                & (dashboard_table$Data_type=='Seasonally adjusted')
                & (dashboard_table$Age_group=='15 years and over'),])
```

The vector for the unemployment rate is `v2062815`, the standard error is 
`v101885192`.




