---
title: "R Notebook"
output: html_notebook
---


# Analysis of Employment - Unemployment Relation

### Goals

* Understand behavior of employment and unemployment
* Understand relation

### Comments

```{r}
library(cansim)
library(stringr)
library(ggplot2)
```

## Retrieval of the individual vectors separately

```{r}
start_date <- "1900-01-01"
```

```{r}
labour_force <- get_cansim_vector('v2062810', start_date)
unemployed <- get_cansim_vector("v2062814", start_date)
unemployment_rate <- get_cansim_vector("v2062815", start_date)
dim(labour_force)
```

All 3 series span the same time range, as expected.
```{r}
summary(as.Date(labour_force$REF_DATE))
summary(as.Date(unemployed$REF_DATE))
summary(as.Date(unemployment_rate$REF_DATE))
```




## Analysis of Time series


```{r}
unemployment_rate$VALUE_MANUAL <- 100.*unemployed$VALUE/labour_force$VALUE
```


Employment reaches new hight, never recorded before! Also, I was able to 
reproduce the unemployment rate from the unemployment numbers and labour force.
```{r}
ggplot(unemployment_rate, aes(x=as.Date(REF_DATE), y=VALUE)) + geom_line() +
  labs(x='reference date', y='unemployment rate') +
  geom_line(aes(x=as.Date(REF_DATE), y=VALUE_MANUAL, color='red'))
```

Interestingly, the labour force, the amount of people who could work, dropped
significantly during Covid-19.
```{r}
ggplot(labour_force, aes(x=as.Date(REF_DATE), y=VALUE)) + geom_line() +
  labs(x='reference date', y='labour force')
```

```{r}
ggplot(unemployed, aes(x=as.Date(REF_DATE), y=VALUE)) + geom_line() +
  labs(x='reference date', y='unemployed persons')
```


