---
title: "R Notebook"
output: html_notebook
---

# Google Trends EDA

### Goals


### Comments


```{r}
suppressMessages(library(dplyr))
suppressMessages(library(purrr))
suppressMessages(library(cansim))
suppressMessages(library(cowplot))
suppressMessages(library(ggplot2))
library(e1071)
library(reshape2)
library(scales)
```


## I. Data Preparation


```{r}
data_path <- "../../../data/GoogleTrends"
google_index_files <- c(
  "multiTimeline_unemployment1_01_07_20.csv",
  "multiTimeline_unemployment2_01_07_20.csv",
  "multiTimeline_unemployment3_01_07_20.csv",
  "multiTimeline_unemployment4_01_07_20.csv",
  "multiTimeline_unemployment5_01_07_20.csv",
  "multiTimeline_unemployment6_01_07_20.csv"
)
```

```{r}
file_contents <- lapply(google_index_files, FUN = function(file) {
  g_group <- read.csv(paste0(data_path, "/", file), skip = 2)
})
g_index <- file_contents %>% reduce(left_join, by = "Month")
```


Retrieve vector with employment numbers: Employment [pers]
```{r}
employment <- get_cansim_vector("v2062809", "1900-01-01")
unemployment_rate <- get_cansim_vector("v2062815", start_date)
dim(employment)
```

```{r}
employment <- select(employment, "REF_DATE", "VALUE")
colnames(employment) <- c("date", "employment")
employment$employment <- employment$employment / 1000. # units in 1M

unemployment_rate <- select(unemployment_rate, "REF_DATE", "VALUE")
colnames(unemployment_rate) <- c("date", "unemployment_rate")
```


```{r}
colnames(g_index) <- lapply(colnames(g_index), function(name) {
  strsplit(name, "\\.\\.\\.")[[1]][1]
})
```


```{r}
g_index_rows <- melt(g_index,
  id = "Month", na.rm = TRUE,
  variable.name = "keyword", value.name = "index"
)
```

```{r}
g_index_rows$year <- sapply(as.character(g_index_rows$Month), FUN = 
                              function(par_date) {
  strsplit(par_date, "-")[[1]][1]
})
g_index_rows$month <- sapply(as.character(g_index_rows$Month), 
                             FUN = function(par_date) {
  strsplit(par_date, "-")[[1]][2]
})
g_index_rows$date <- as.Date(paste(g_index_rows$Month, "-01", sep = ""))
```


```{r}
g_index_rows$index_num <- as.numeric(g_index_rows$index)
g_index_rows[is.na(g_index_rows$index_num), "index_num"] <- 0.
g_index_rows$index <- g_index_rows$index_num
g_index_rows <- g_index_rows[order(g_index_rows$date), ]
```

```{r}
unemployment_rate <- filter(unemployment_rate, date >= as.Date("2004-01-01"))
```

## II. Modeling with unemployment keyword time series

I pick the time series based on the frequency of the 'unemployment' keyword
for further anlaysis.

* Linear Regression model sensitive to outliers, such as those from the recession.
* Effect of non-stationarity on regression models: non-statinary time series
can lead to spurious regression.


Definition of the recession 08/09 and the Covid-19 related recession starting
in May 2020.
```{r}
recession_years <- as.Date(c(
  "2008-01-01", "2009-12-31", "2020-04-01",
  "2020-06-01"
))
```


Limit the data to only include time points until including 2020-05-01.
```{r}
g_index_unempl <- filter(
  g_index_rows, keyword == "unemployment",
  date < as.Date("2020-06-01")
)
unemployment_rate <- filter(
  unemployment_rate,
  date < as.Date("2020-06-01")
)
```



```{r}
unemployment_rate$uer_log <- log(unemployment_rate$unemployment_rate)
```




Test with seasonal AR(1) model, as Model 0.0 in Tuhkuri16

```{r}
model0 <- Arima(unemployment_rate$uer_log, order = c(1, 0, 0), seasonal = c(1, 0, 0))
model0
```

There is still some autocorrelation in the residuals, at lags 1 and 2 but
disappears at higher lag. As
in Tuhkuri16  A.2 I have PACF has two lags beyond above 5%.
```{r}
model0 %>%
  residuals() %>%
  ggtsdisplay()
```

To check for autocorrelation in the residuals I use the Ljung–Box test statistic. We
can reject that H0 hypothesis that the data comes from a white noise distribution and
hence there is still autocorrelation present..
```{r}
model0 %>%
  residuals() %>%
  Box.test(lag = 12, fitdf = 1, type = "Lj")
```

```{r}
length(unemployment_rate$unemployment_rate)
length(g_index_unempl$index)
```

Model 1.0
```{r}
model1 <- Arima(unemployment_rate$uer_log,
  order = c(1, 0, 0), seasonal = c(1, 0, 0),
  xreg = g_index_unempl$index, method = "ML"
)
model1
```

```{r}
checkresiduals(model1)
```

```{r}
fcast <- forecast(model1, xreg = g_index_unempl$index, h)
```


```{r}
accuracy(fcast, unemployment_rate$uer_log)
```