---
title: "TestSetup"
output: html_notebook
# output: rmarkdown::github_document
# output: 
#  html_document:
#    keep_md: true
---

# Exploratory analysis of series with a potential to nowcast GDP

### Goals

* Plot series used in the Index of the Canadian Economic Dashboard and COVID-19
* Identify data tables and vectors for the dashboard

### Comments

* Time series from Canadian Economic Dashboard and COVID-19 at https://www150.statcan.gc.ca/n1/pub/71-607-x/71-607-x2020009-eng.htm
* There are 16 indicators


```{r}
suppressMessages(library(cansim))
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(plotly))
source("../../src/utils.R")
```

## Extract the Data
```{r}
vector_names <- c(
  "v65201210", "v41690973", "v2062809", "v1001827265",
  "v1001826653", "v52367097", "v4391505", "v800450",
  "v32858858", "v32858865", "v32858872", "v74869", "v129449",
  "v129460", "v129472", "v129482"
)
length(vector_names)
```

```{r}
vector_descr <- c(
  "RealGDP", "CPI", "Employment [pers]", "Int. merchandise trade Exp. [$]",
  "Int. merchandise trade Imp. [$]", "Retail Sales [$]", "hours worked",
  "Manufact. sales [$]", "Aircraft domestic [#]", "Aircraft transborder [#]",
  "Aircraft int other [#]", "Railway carloads [tons]", "Travelers US [pers]",
  "Travelers other country [pers]", "CA resident US [pers]",
  "CAresident other country [pers]"
)
names(vector_descr) <- vector_names
```

I expect this date to be before the start of all time series/vectors.
```{r}
start_date <- "1980-01-01"
```

Retrieve all vectors. They are concatenated along axis 0.
```{r}
vectors <- get_cansim_vector(vector_names, start_date)
dim(vectors)
```


## Preprocessing


```{r}
colnames(vectors) <- sapply(colnames(vectors), tolower)
colnames(vectors)
```

```{r}
vectors$ref_date <- as.Date(vectors$ref_date)
```


```{r}
head(vectors)
```

```{r}
vector_names[1]
```

```{r}
length(vector_names)
```


```{r}
indicators1 <- vector(length = 16)
for (i in 1:length(vector_names)) {
  indicator <- filter(vectors, vector == vector_names[i])
  ref_value <- indicator$value[indicator$ref_date == as.Date("2019-12-01")]
  indicator$value_index <- 100. * indicator$value / ref_value
  if (i == 1) {
    indicators <- indicator
    }
  else {
    indicators <- rbind(indicators, indicator)
    }
}
```

```{r}
head(indicators)
```

```{r}
indicators$Descr <- sapply(indicators$vector, function(x) {
  vector_descr[x][[1]]
})
```




## Canadian Economic Dashboard Visualization

```{r dashboard, fig.height=5,fig.width=11}
p <- ggplot(indicators, aes(x = ref_date, y = value_index)) +
  geom_line((aes(group = Descr, color = Descr))) +
  scale_x_date(date_breaks = "3 month", date_labels = "%b %Y",
               limits = as.Date(c("2019-01-01", "2020-09-01"))) +
  scale_y_continuous(breaks = seq(0, 180, 20), limits = c(0, 180)) +
  labs(title = "Index, December 2019 = 100", x = "", y = "")
ggplotly(p)
```
## Convert Dashboard series into a tsibble for forecasting

* Start Date = January, 2000

```{r, echo=FALSE}
filter(indicators,Descr=="RealGDP" & ref_date >= "2000-01-01")%>%select(value,ref_date)%>%mutate(RealGDP=value, Month=yearmonth(ref_date))%>%select(Month,RealGDP)->x1
filter(indicators,Descr=="CPI" & ref_date >= "2000-01-01")%>%select(value,ref_date)%>%mutate(CPI=value, Month=yearmonth(ref_date))%>%select(Month,CPI)->x2
dashboard<-full_join(x1,x2,by="Month")
filter(indicators,Descr=="Int. merchandise trade Exp. [$]" & ref_date >= "2000-01-01")%>%select(value,ref_date)%>%mutate(Export=value, Month=yearmonth(ref_date))%>%select(Month,Export)->x3
dashboard<-full_join(dashboard,x3,by="Month")
filter(indicators,Descr=="Int. merchandise trade Imp. [$]" & ref_date >= "2000-01-01")%>%select(value,ref_date)%>%mutate(Import=value, Month=yearmonth(ref_date))%>%select(Month,Import)->x4
dashboard<-full_join(dashboard,x4,by="Month")
filter(indicators,Descr=="Retail Sales [$]" & ref_date >= "2000-01-01")%>%select(value,ref_date)%>%mutate(Retail=value, Month=yearmonth(ref_date))%>%select(Month,Retail)->x5
dashboard<-full_join(dashboard,x5,by="Month")
filter(indicators,Descr=="hours worked" & ref_date >= "2000-01-01")%>%select(value,ref_date)%>%mutate(Hours=value, Month=yearmonth(ref_date))%>%select(Month,Hours)->x6
dashboard<-full_join(dashboard,x6,by="Month")
filter(indicators,Descr=="Manufact. sales [$]" & ref_date >= "2000-01-01")%>%select(value,ref_date)%>%mutate(Manuf=value, Month=yearmonth(ref_date))%>%select(Month,Manuf)->x7
dashboard<-full_join(dashboard,x7,by="Month")
filter(indicators,Descr=="Railway carloads [tons]" & ref_date >= "2000-01-01")%>%select(value,ref_date)%>%mutate(Rail=value, Month=yearmonth(ref_date))%>%select(Month,Rail)->x8
dashboard<-full_join(dashboard,x8,by="Month")
dashboard<-as_tsibble(dashboard,index=Month)
rm(x1,x2,x3,x4,x5,x6,x7,x8,i)
```

```{r}
head(dashboard, echo=TRUE)
```




## Plot series

```{r Forecasting Series,fig.height=5,fig.width=11}
as_tsibble(filter(indicators, ref_date >= "2000-01-01")%>%select(value,ref_date,Descr)%>%mutate(Month=yearmonth(ref_date),Series=Descr)%>%select(Month,value,Series),index=Month,key=Series)%>%mutate(Date=as.Date(Month))->SeriesData1
p <- ggplot(SeriesData1, aes(x = Date, y = value)) +
     geom_line((aes(group = Series, color = Series))) +
     scale_x_date(date_breaks = "2 year", date_labels = "%Y",
                  limits = as.Date(c("2000-01-01", "2020-09-01"))) +
     labs(title = "Series for Forecasting", x = "", y = "")
ggplotly(p)
```