---
title: "R Notebook"
output: html_notebook
---

# Test R Package Quantmod

### Goals

* Test functionality of quantmod package to retrive data

### Comments

* Focus on BPY	Brookfield Property Partners L.P. suggested by David (SM)
* There is a difference between BPY (NasdaqGS) and BPY-UN.TO 
* Each stock has multiple values (e.g. open, close). Which ones are the most valuable to include in our models?
* Questions:
  * comparison with Yahoo displayed data, is it open/close?
  * How to interpret the tickers?BPY-UN.TO


```{r}
rm(list=ls())
library(quantmod)
```


Manual acces to 
https://finance.yahoo.com/quote/BPY?p=BPY&.tsrc=fin-srch


```{r}
bpy_stock <- getSymbols("BPY", src = "yahoo", auto.assign = FALSE) 
```


columns Open, High, Low, Close, Volume, and Adjusted (Yahoo): 

```{r}
head(bpy_stock)
```

```{r}
class(bpy_stock)
```

```{r}
min(index(bpy_stock))
max(index(bpy_stock))
```

```{r}
plot_close <- function (stock) {
  # plots daily close stock price
  stock_name <- str_split(colnames(stock)[1], "\\.", simplify = TRUE)[1]
  ggplot(stock, aes(x = index(stock), y = stock[, 4])) + geom_line(color = "darkblue") + 
      ggtitle(paste(stock_name, "Close Value")) + xlab("Date") + 
    ylab("Price") + theme(plot.title = element_text(hjust = 0.5)) + 
    scale_x_date(date_labels = "%b %y", date_breaks = "6 months") 
}
```

```{r}
plot_close(bpy_stock)
```



### BPY-UN.TO 

Data for Toronto TSX at https://ca.finance.yahoo.com/quote/BPY-UN.TO/history?p=BPY-UN.TO

```{r}
bpy_tsx_stock <- getSymbols("BPY-UN.TO", src = "yahoo", auto.assign = FALSE) 
```

```{r}
head(bpy_tsx_stock)
```

```{r}
min(index(bpy_tsx_stock))
max(index(bpy_tsx_stock))
```

```{r}
plot_close(bpy_tsx_stock)
```


### Open-to-close Percentage change


```{r}
head(OpCl(bpy_tsx_stock))
```


```{r}
summary(bpy_tsx_stock)
```


Stock returns: https://financeformulas.net/Total-Stock-Return.html

```{r}
head(monthlyReturn(bpy_tsx_stock))
```

## Extracting Enterprise Value

Enterprise Value = market cap + debt

```{r}
metrics <- yahooQF(c("Name", "Enterprise Value", "Enterprise Value/Revenue", "Enterprise Value/EBITDA", "Total Cash Per Share (mrq)", "EBITDA", "Volume", "P/E Ratio", "Dividend Yield", "Shares Outstanding", "Price/Book"))
```

```{r}
getQuote("BPY", what = yahooQF(c("Name", "Enterprise Value")))
```

```{r}
head(stats)
```



```{r}
metrics <- yahooQF(c("Name", "Enterprise Value", "Enterprise Value/Revenue", "Enterprise Value/EBITDA", "Total Cash Per Share (mrq)", "EBITDA", "Volume", "P/E Ratio", "Dividend Yield", "Shares Outstanding", "Price/Book"))

stats <- getQuote(symbols, what = metrics) %>%
  rownames_to_column("symbol")
```



```{r}
ap_tsx_stock <- getSymbols("AP-UN.TO", src = "yahoo", auto.assign = FALSE) 
```


```{r}
head(ap_tsx_stock)
```

```{r}
head(monthlyReturn(ap_tsx_stock))
```

```{r}
stock = ap_tsx_stock
stock = monthlyReturn(stock)
colnames(stock) <- sapply(colnames(stock), function (colstock) str_split(colstock, "\\.")[[1]][length(str_split(colstock, "\\.")[[1]])])
stock <- data.frame(date=index(stock), coredata(stock))
stock
```



```{r}
class(bpy_tsx_stock)
```

```{r}
head(monthlyReturn(ap_tsx_stock$`AP-UN.TO.Close`))
```


